---
title: "derivative"
author: "Qi Yuchen, yq2279"
date: "2020/4/23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stats)
library(matrixcalc)
library(pracma)
library(qpcR)
library(pspline)
library(patchwork)
```

# data import and cleaning

```{r}
df.raw = read.csv("covid19-1.csv")
df = df.raw %>% 
  janitor::clean_names() %>% 
  dplyr::select(country_region, province_state, date, confirmed_cases) %>% 
  filter(confirmed_cases != 0)

region.index = levels(df$country_region)

# first step to write a function
df.china = df %>% 
  filter(country_region == region.index[30]) %>% 
  group_by(country_region, date) %>% 
  summarise(cases = sum(confirmed_cases)) %>% 
  mutate(formal_date = as.numeric(as.character(as.Date(date, '%m/%d/%Y'), format="%Y%m%d"))) %>% 
  mutate(time = formal_date-min(formal_date)) %>% 
  dplyr::select(region = country_region, date, time, cases) %>% 
  arrange(time)

# function to get data for a specific region
df.region = function(df, region) {
  df.r = df %>% 
    filter(country_region == region) %>% 
    group_by(country_region, date) %>% 
    summarise(cases = sum(confirmed_cases)) %>% 
    mutate(formal_date = as.numeric(as.character(as.Date(date, '%m/%d/%Y'), format="%Y%m%d"))) %>% 
    mutate(time = formal_date-min(formal_date)) %>% 
    dplyr::select(region = country_region, date, time, cases) %>% 
    arrange(time)
  
  df.r
}


df.china = df.region(df, region.index[30])
df.us = df.region(df, "US")

df.test = df.region(df, "US")
test.result = adam(df.test,theta.initial = c(100000,	0.1, 200),maxt = 1e+6)
test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+6,alpha = 0.001)
test.result

# China
df.test = df.region(df, "China")
nls.china = nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 80000,b=0.0869951070488976,c=56.0118879117201))
nls.result = summary(nls.china)$parameters[,1]

adam.result = adam(df.test,theta.initial =nls.result,maxt = 1e+6)
adam.result
adam2 = adam(df.test,theta.initial =c(80000, 2.186193e-01, 90),maxt = 1e+6)

yplot.adam = f.logit(adam2[2:4],t = df.test$time)
plot(df.test$time,yplot.adam, ylim=c(0,90000))
par(new=TRUE)
plot(df.test$time,df.test$cases,col="green",  ylim=c(0,90000))
```

adam                 a            b            c
[1] t=1.000000e+06 1.019762e+05 3.986966e-01 9.977302e+01 alpha=0.002
[2] t=1.000000e+06 1.029744e+05 3.971065e-01 9.981956e+01 alpha=0.001

nls           a            b            c 
1.130883e+05 3.830366e-01 1.002671e+02 


china nls          a            b            c 
         7.867403e+04 2.186193e-01 8.685954e+01 

# optimization

## gradient descent

```{r newton gradient}
f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

# objective function to maximize
obj = function(beta, x, y) {
  a = beta[1]
  b = beta[2]
  c = beta[3]
  
  y.fit = f.logit(beta, x)
  
  2*dot(y, y.fit) - dot(y.fit, y.fit)
}


# objective, gradient and hessian
logisticstuff <- function(x, y, beta) {
  # obj
  obj = obj(beta, x, y)
  # gradient
  f = function(beta) {
      a = beta[1]
      b = beta[2]
      c = beta[3]
      
      y.fit = f.logit(beta, x)
  
      2*dot(y, y.fit) - dot(y.fit, y.fit)
  }
  grad = grad(f, beta)  
  # Hessian
  hess = -diag(3)
 
  return(list(obj = obj, grad = grad, hess = hess))
}

# function to get the optimal beta (maximize)
logitcurve = function(X, Y, func = logisticstuff, start, tol = 1e-5, maxiter = 1000) {
  i <- 0
  cur <- start
  stuff <- func(X, Y, cur)
  res <- c(0, stuff$obj, cur)
  prevobj <- -Inf
  prev = Inf
  while(i < maxiter && abs(stuff$obj - prevobj) > tol) {
    i <- i + 1
    prevobj <- stuff$obj
    prev <- cur
    
    hess = stuff$hess
    grad = stuff$grad
    while (!is.negative.definite(hess)) {
      hess = hess - diag(start)
    }
    
    cur <- prev - solve(hess) %*% grad
    stuff <- func(X, Y, cur) # obj, gradient, Hessian
    # step-halfing
    l=1
    while(stuff$obj <= prevobj && l < 1000){
      cur <- prev - (1/(2^l))*solve(hess) %*% grad
      stuff = func(X, Y, cur)
      l= l+1
    }
     res <- rbind(res, c(i, stuff$obj, cur))
  }
  
  return(res)
}

# test
df.test = df.region(df, "Romania")
test.result = logitcurve(df.test$time, df.test$cases,start = c(1599.997,	0.2641970,	98.94560))

test.result2 = logitcurve(df.test$time, df.test$cases,start = c(1.936561e+06, 0.19, 1.376097e+02))
```


## adam

```{r adam}
f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}
f = function(beta, x, y) {
  a = beta[1]
  b = beta[2]
  c = beta[3]
  
  y.fit = f.logit(beta, x)
  
  - 2*sum(y * y.fit) + sum(y.fit * y.fit)
}
f.grad = function(beta, x, y) {
  a = beta[1]
  b = beta[2]
  c = beta[3]
  
  a.grad = sum(2*a/(1+exp(-b*(x-c)))^2-2*y/(1+exp(-b*(x-c))))
  b.grad = sum(2*a^2*(x-c)*exp(-b*(x-c))/(1+exp(-b*(x-c)))^3+2*a*(c-x)*y*exp(-b*(x-c))/(1+exp(-b*(x-c)))^2)
  c.grad = sum(-2*a^2*b*exp(-b*(x-c))/(1+exp(-b*(x-c)))^3+2*a*b*y*exp(-b*(x-c))/(1+exp(-b*(x-c)))^2)
  theta = c(a.grad,b.grad,c.grad)
  return(theta)
}


adam = function(data,alpha = 0.002, beta1 = 0.9, beta2 = 0.999, theta.initial, epsilon = 1e-8, tol = 1e-10, maxt = 10000) {
  x = data$time
  y = data$cases
  m.cur = 0
  v.cur = 0
  t = 0
  theta.prev = Inf
  theta.cur = theta.initial
  while(abs(mean(theta.cur-theta.prev))>tol&& t<maxt) {
    theta.prev = theta.cur
    m.prev = m.cur
    v.prev = v.cur
    t = t+1
    g = f.grad(theta.prev, x,y)
    m.cur = beta1*m.prev + (1-beta1)*g
    v.cur = beta2*v.prev + (1-beta2)*g^2
    mbar = m.cur/(1-beta1^t)
    vbar = v.cur/(1-beta2^t)
    theta.cur = theta.prev - alpha*mbar/(sqrt(vbar)+epsilon)
  }
  c(t,theta.cur)
}

# test
df.test = df.region(df, "Vietnam")
test.result = adam(df.test,theta.initial = c(300,	0.1, 200),maxt = 1e+5)
test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+6,alpha = 0.002)
test.result

f(test.result[2:4],df.test$time,df.test$time) - f(c(780.800001 ,  0.111383 ,215.153323 ),df.test$time,df.test$time)


  
```
> # test
> df.test = df.region(df, "Vietnam")
> test.result = adam(df.test,theta.initial = c(0,	0, 0),maxt = 1e+5)
> test.result
[1] 1.000000e+05 1.163637e+02 2.309543e-01 1.912774e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 3.062653e+02 1.314710e-01 2.033377e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 4.442966e+02 1.204103e-01 2.081498e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 5.075830e+02 1.177624e-01 2.098387e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 5.460322e+02 1.163397e-01 2.107562e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 5.734413e+02 1.155234e-01 2.113679e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 5.946681e+02 1.149577e-01 2.118197e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 6.118883e+02 1.145170e-01 2.121737e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 6.262789e+02 1.141760e-01 2.124610e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5)
> test.result
[1] 1.000000e+05 6.385854e+02 1.138981e-01 2.127010e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.005)
> test.result
[1] 1.000000e+05 6.486948e+02 1.136925e-01 2.128920e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.01)
> test.result
[1] 1.000000e+05 6.553686e+02 1.135459e-01 2.130186e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.01)
> test.result
[1] 1.000000e+05 6.612772e+02 1.134874e-01 2.131246e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.01)
> test.result
[1] 1.000000e+05 6.665094e+02 1.133739e-01 2.132223e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.002)
> test.result
[1] 1.000000e+05 6.740703e+02 1.131599e-01 2.133656e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.002)
> test.result
[1] 1.000000e+05 6.809052e+02 1.130233e-01 2.134889e+02
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+5,alpha = 0.002)
> test.result
[1] 1.000000e+05 6.871018e+02 1.129101e-01 2.136000e+02
> f(test.result[2:4],df.test$time,df.test$time) - f(c(780.800001 ,  0.111383 ,215.153323 ),df.test$time,df.test$time)
[1] 170.442
> test.result = adam(df.test,theta.initial = test.result[2:4],maxt = 1e+6,alpha = 0.002)
> test.result
[1] 1.000000e+06 7.276823e+02 1.121954e-01 2.142998e+0

# analyze

 [1] "Qatar"                "Romania"              "Russia"               "Rwanda"               "San Marino"           "Saudi Arabia"         "Senegal"              "Serbia"               "Singapore"           
[10] "Slovakia"             "Slovenia"             "South Africa"         "Spain"                "Sri Lanka"            "Sweden"               "Switzerland"          "Taiwan*"              "Thailand"            
[19] "Trinidad and Tobago"  "Tunisia"              "Turkey"               "Ukraine"              "United Arab Emirates" "United Kingdom"       "Uruguay"              "US"                   "Uzbekistan"          
[28] "Venezuela"            "Vietnam" 

```{r}

```


## Vietnam

```{r}
df.test = df.region(df, region.index[162])

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 140)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 140)
p1 + p2
plot(x = df.test$cases, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(5.999722796,6.132399253) # from f1
pred = c(91,134) # from df.test$cases
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 99.5
beta = c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(450,0.1,210), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) + ylim(0, 400) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5) + ylim(0, 400)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 450, b = 0.1, c = 210))
nls.result = summary(nls.fit)$parameters[,1]
```
         a          b          c 
780.800001   0.111383 215.153323 


## Venezuela 

hard to fit, almost linear trend

```{r}
df.test = df.region(df, region.index[161])

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 140)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 140)
p1 + p2
plot(x = df.test$cases, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(f1[10],f1[11]) # from f1
pred = c(df.test$cases[10],df.test$cases[11]) # from df.test$cases
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 99.5
beta = c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(95.28,0.426,5.48), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) + ylim(0, 400) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5) + ylim(0, 400)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 2000, b = 0.3, c = 20))
nls.result = summary(nls.fit)$parameters[,1]
```
         a          b          c 
95.2801029  0.4263547  5.4823632 

## Uzbekistan

```{r}
df.test = df.region(df, region.index[160])

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 50)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 50)
p1 + p2
plot(x = f0, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(f1[9],f1[10]) # from f1
pred = c(f0[9],f1[10]) # from df.test$cases
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 6
beta = c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(50,1.1,5), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) + ylim(0, 50) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5) + ylim(0, 50)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 50, b = 1.1, c = 5))
summary(nls.fit)$parameters[,1]
ybar = f.logit(summary(nls.fit)$parameters[,1], df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) + ylim(0, 50) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5) + ylim(0, 50)
p1+p3
```
         a          b          c 
50.1234967  0.7303043  4.0919766 

## Uruguay

```{r}
df.test = df.region(df, region.index[158])

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 200)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 200)
p1 + p2
plot(x = f0, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(f1[9],f1[10]) # from f1
pred = c(f0[9],f1[10]) # from f0
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 6
beta = c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(154,1.1,6), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) + ylim(0, 200) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5) + ylim(0, 200)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 154, b = 1.1, c = 6))
summary(nls.fit)$parameters[,1]
ybar = f.logit(summary(nls.fit)$parameters[,1], df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5)
p1+p3
```
         a          b          c 
173.5742876   0.5981794   5.6981907

## Qatar

```{r}
df.test = df.region(df, "Qatar")

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 600)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 600)
p1 + p2
plot(x = f0, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(f1[23],f1[14]) # from f1
pred = c(f0[23],f1[14]) # from f0
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 6
c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(700,1,90), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 1) + ylim(0, 600) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 1) + ylim(0, 600)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a = 700, b = 1, c = 90))
summary(nls.fit)$parameters[,1]
ybar = f.logit(summary(nls.fit)$parameters[,1], df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5)
p1+p3
```
         a          b          c 
480.403252   0.637724  83.029500 

## Romania

```{r}
df.test = df.region(df, "Romania")

p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point() + ylim(0, 800)
smooth.fit = smooth.Pspline(x = df.test$time, y = df.test$cases,df=5, method=2)

f0 = predict(smooth.fit, df.test$time, nderiv=0)
f1 = predict(smooth.fit, df.test$time, nderiv=1)
p2 =  ggplot(mapping = aes(x = df.test$time, y = f0)) + geom_point() + ylim(0, 800)
p1 + p2
plot(x = f0, y = f1)

# select two points to fit a polynomial degree = 2 and through (0,0)
resp = c(f1[21],f1[20]) # from f1
pred = c(f0[21],f1[20]) # from f0
y1 = pred[1];y2 = pred[2]
y1d = resp[1]; y2d = resp[2]
a = (y2d*y1^2 - y1d*y2^2) / (y2d*y1 - y1d*y2)
b = y1d/(y1*(1-y1/a))
c = 6
c(a,b,c)

f.logit=function(beta, t){
  a = beta[1]
  b = beta[2]
  c = beta[3]
  a/(1+exp(-b*(t-c)))
}

ybar = f.logit(c(1.936561e+06, 0.19, 1.376097e+02), df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 1) + ylim(0, 800) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 1) + ylim(0, 800)
p1+p3
```

```{r}
# use nls
nls.fit <- nls(cases ~ a/(1+exp(-b*(time-c))), df.test, start = list(a=162, b=	0.4,c=	13),control = list(maxiter = 10000,minFactor=1/2621446666666,warnOnly = TRUE))
summary(nls.fit)$parameters[,1]
ybar = f.logit(summary(nls.fit)$parameters[,1], df.test$time)
p1 = ggplot(mapping = aes(x = df.test$time, y = df.test$cases)) + geom_point(size = 0.5) 
p3 = ggplot(mapping = aes(x = df.test$time, y = ybar)) + geom_point(size = 0.5)
p1+p3
```

